### 2021-08-15 doc/thesis/chapters/implementation.tex
sample dec
all /ref to ~/ref (to avoid newline separations)

The theory behind the demodulation of an FM signal is already explained in Chapter \ref{cha:SignalProcessingTheory}.
The demodulation of an FM broadcast channel specifically, is described in Section \ref{sec:demodulation-of-broadcast-fm}.
In this section however, the DSP chain is described into more detail and with more information, as it is implemented in the various methods.\\

\noindent
Figure \ref{fig:bd_dsp_detailed} displays the detailed DSP chain as a block diagram.\\

@@

  \item \textbf{FM Demodulator}\\
       TODO\\
       % TODO
      The FM demodulator is implemented as a Baseband Delay Modulator, as described in Section \ref{sec:BasebandDelayDemodulator}.

@@

        \item \textbf{Between the branches \textit{Recover LR diff} and \textit{Recover Carriers}:}\\
          Both branches only contain a single BP filter.
          However, they have properties, like the transition bandwidth and the passband.
          Thus, their filter order may be different as well, which leads to a non-matching group delay.
          Since the two signals are multiplicated in the subsequent mixer, and an important factor is the phase-coherency of the recovered carrier (as explained in Section \ref{sec:demodulation-of-broadcast-fm}), the group delay compensation becomes necessary here.\\

@@

  \item \textbf{Carrier Recovery}\\
      The 38~kHz carrier can be recovered from the 19~kHz pilot tone.
      It is important that the recovered carrier is phase-coherent with the received input, as described in Section \ref{sec:demodulation-of-broadcast-fm}.
      This can be achieved, by simply multiplying the pilot tone with itself, which is the way it is implemented here.\\
\end{itemize}

@@

The transmitter is implemented according to the structure that is depicted as a block diagram in Figure \ref{fig:bd_matlab_transmitter}.
To facilitate an easy understanding of the block diagram, it is of advantage to keep in mind that it exactly resembles the frequency spectrum that is previously shown in Figure \ref{fig:fm_channel_baseband_freqs}.\\

@@

The achieved sample rate of 960~kHz could actually be chosen much lower, since an FM channels' maximum frequency is 75~kHz (see Section \ref{sec:FrequencySpectrumOfABroadcastFmChannel}), and according to the Nyquist theorem the sample rate thus only needs to have 150~kHz (see Section \ref{sec:SamplingTheorem}).
However, the higher rate was chosen, mainly because a higher number of signal samples beautify any plots that are created during analysis.
Since a thorough optimization of the DSP chain is not the focus of this thesis, this was an acceptable trade-off.\\

@@

At this point in the DSP chain, the signal resembles the frequency spectrum that is displayed in Figure \ref{fig:fm_channel_baseband_freqs}.\\

@@

The receiver is implemented, following the block diagram in Figure \ref{fig:bd_dsp_detailed}.
Further assumptions and simplifications, such as the fixed point arithmetic are described into detail above.

@@

An example of a plot that is created by the analysis is depicted in Figure \ref{fig:matlab_analysis_time_domain}.
The diagram shows the respective signals in the time domain.
Above two timelines display the input data, as it is read from the input audio file.
The two diagrams in the center plot the Mono and LR Diff signals.
The bottom diagrams then represent the result after the entire signal processing chain.
It can be seen that the DSP chain is not able to demodulate the original signal without any error.
However, the result shows a clear separation of the left and right channels, as resembles the original input very closely, which is good enough for this case study.\\

@@

Further analysis is done in the frequency domain, especially in order to better understand the modulation of the different spectral parts in the FM channel.
An example thereof is shown in Figure \ref{fig:matlab_analysis_freq_domain}.
The spectrum of an FM channel multiplex signal, as it is generated by a transmitter before the modulation, is shown in this diagram.
The spectrum after the demodulation through a receiver is shown in the lower diagram.
Spectral parts, such as the pilot tone at 19~kHz, or the LR Diff part around 38~kHz can be seen clearly.
The demodulated spectrum shows an artifact at 57~kHz, which does not exist in the original transmitter spectrum.
This is an artifact of the demodulation process, specifically the frequency-shift of the LR Diff part to baseband.
The artifact is very low in energy, because it is attenuated by a BP filter in the down-conversion process, and can thus be ignored for further processing, even though it can still be seen in the spectral analysis.

@@

GNU Radio provides a comprehensive set of DSP blocks, which are available to build various functionalities.
Multiple blocks are connected to a block diagram, that is very similar to the one in Figure \ref{fig:bd_matlab_transmitter}.
The USRP device is used to broadcast the RF signal over the air.\\

@@

Special care needs to be taken when transmitting signals over the air.
The frequency spectrum is strongly limited by the radio regulations and may be monitored by the authorities.
Please make sure to adapt your setup to the respective local regulations, to avoid any legal issues.
Some of the regulations are mentioned in Section \ref{sec:FrequencyBand}, but may differ in various countries, however.\\

\noindent
Figure \ref{fig:bd_gnuradio_setup_transmitter} shows a block diagram of the transmitter application.

@@

This receiver works with both devices, the USRP b200mini and the RTL-SDR.
Figure \ref{fig:bd_gnuradio_setup_receiver} shows a block diagram of the receiver application.

@@

Figure \ref{fig:bd_cocotb_interface_simulator} gives an overview of how cocotb interfaces with a simulator.
The Python test cases are shown as a list of coroutines, which are managed by a scheduler.
This Python code interfaces with the simulator through standard interfaces like VPI.
The block called \textit{GPI} is developed by cocotb and abstracts the different simulator interfaces, such as VPI, VHPI or FLI.
In the simulator, the device under test (DUT) is instantiated.
The RTL simulation itself is managed by the simulators' scheduler.

@@

The architecture for this testbench is created under the assumptions and according to the guidelines that are developed in Chapter \ref{cha:SystemArchitectureAndConcept}.\\

Figure \ref{fig:bd_testbench_architecture_vhdl} shows a block diagram of the implemented architecture.
The split between the Python environment and the VHDL implementation is clearly visible but is represented from the developers perspective.
The actual simulator interfaces, as described in Section \ref{sub:VHDL:FrameworkCocotb}, are hidden to maintain the overview of the implemented architecture.\\

@@

Figure \ref{fig:register_engine} gives an overview of how the Register Engine script works.
The Python script takes the register definitions from the YAML file, and uses templates to generate the respecitive output files.\\

@@

A testbench is equally important in the HLS implementation, as it is for the VHDL implementation.
The HDL code for the IP is written in \cplusplus\ and thus the testbench is also written in this language.
The execution of the testbench is done in the Xilinx HLS software.
A detailed explanation of how an HLS testbench works in general is given in Chapter \ref{cha:HLS}.\\

@@

The testbench is especially designed around the 'Design For Reuse' segment of the SIMILAR concept, which is introduced in Section \ref{sec:SimilarProcess}.
This is the case because of two major parts.
%It is not only using the exact same Python code for analysis, but is also using parts of the application software code directly, to load the testbench input data.\\

@@

In the previous Section \ref{sec:impl:hls:testbench} it is explained, that the HLS testbench generates output files during its execution.
These output files are analyzed with the exact same Python analysis code, that is already used in the VHDL testbench analysis.

@@

The testbench is developed in the language \cplusplus\ and does not have any restrictions in the usage thereof, as Section \ref{sec:hls:workflow} described.
Thus, the \cplusplus\ code that is later used in the application software to interface with the IP, can already be used in the testbench here, for the same purpose.
This is especially useful in the development process of such a combined system, consisting of software and hardware, that relies on these two parts working together successfully.
The fact that the application software can directly be used in the testbench, enables the development of software that interfaces with the hardware in a very early stage of the project, before any actual hardware exists.
It also allows that the software part is already tested before the deployment on hardware.\\

Figure \ref{fig:testbench_hls_overview_reuse} gives an overview of the testbench and specifically points out the reuse of the application software.
The diagram displays the HLS IP in the testbench and in the hardware with dotted lines, to emphasize that they are actually the same.\\

@@

HLS can automatically implement interfaces, as described in Section \ref{sec:hls:coding:interfaces}.
This is a major feature that is used in the implementation of the HLS IP.\\

The design uses an AXI4-Lite memory mapped interface, as well as multiple AXI4-Streams.
Additionally, a data-only interface is used for the connected LEDs.
The actual implementation is demonstrated in Program \ref{prog:hls_impl_interfaces}.
These few lines of code already implement the entire functionality of each interface type.\\

@@

%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Sample Rate Reduction}

Multiple steps in the DSP chain require the operation of sample rate reduction.
This is explained in Section \ref{sec:impl:dsp-chain}.\\

@@

The implementation of the sample rate reduction can be shown in a short code example in Program \ref{prog:sample_rate_reduction_fifo_behaviour}.
The FM Demodulator runs in the oversampled domain and is executed \texttt{OSR\_RX} times in the inner for-loop.
Only the last sample is written to the \texttt{fm\_channel\_data} variable, which itself is an AXI stream type with an underlying FIFO depth of \texttt{OSR\_AUDIO}, as defined with the accompanying pragma.
Thus, the inner for-loop is executed \texttt{OSR\_AUDIO} times by the outer loop, to fill this FIFO until it is full.
The subsequent Channel Decoder then reads the entire FIFO and also implements downsampling with a similar behaviour.
Several samples are processed, where only the last one is forwarded to the subsequent processing step.
The FIFO behaviour of AXI stream variables is described in Section \ref{sec:hls:coding:interfaces}.

>>>>

The theory behind the demodulation of an FM signal is already explained in Chapter~\ref{cha:SignalProcessingTheory}.
The demodulation of an FM broadcast channel specifically, is described in Section~\ref{sec:demodulation-of-broadcast-fm}.
In this section however, the DSP chain is described into more detail and with more information, as it is implemented in the various methods.\\

\noindent
Figure~\ref{fig:bd_dsp_detailed} displays the detailed DSP chain as a block diagram.\\

@@

  \item \textbf{FM Demodulator}\\
       TODO\\
       % TODO
      The FM demodulator is implemented as a Baseband Delay Modulator, as described in Section~\ref{sec:BasebandDelayDemodulator}.

@@

        \item \textbf{Between the branches \textit{Recover LR diff} and \textit{Recover Carriers}:}\\
          Both branches only contain a single BP filter.
          However, they have properties, like the transition bandwidth and the passband.
          Thus, their filter order may be different as well, which leads to a non-matching group delay.
          Since the two signals are multiplicated in the subsequent mixer, and an important factor is the phase-coherency of the recovered carrier (as explained in Section~\ref{sec:demodulation-of-broadcast-fm}), the group delay compensation becomes necessary here.\\

@@

  \item \textbf{Carrier Recovery}\\
      The 38~kHz carrier can be recovered from the 19~kHz pilot tone.
      It is important that the recovered carrier is phase-coherent with the received input, as described in Section~\ref{sec:demodulation-of-broadcast-fm}.
      This can be achieved, by simply multiplying the pilot tone with itself, which is the way it is implemented here.\\
\end{itemize}

@@

The transmitter is implemented according to the structure that is depicted as a block diagram in Figure~\ref{fig:bd_matlab_transmitter}.
To facilitate an easy understanding of the block diagram, it is of advantage to keep in mind that it exactly resembles the frequency spectrum that is previously shown in Figure~\ref{fig:fm_channel_baseband_freqs}.\\

@@

The achieved sample rate of 960~kHz could actually be chosen much lower, since an FM channels' maximum frequency is 75~kHz (see Section~\ref{sec:FrequencySpectrumOfABroadcastFmChannel}), and according to the Nyquist theorem the sample rate thus only needs to have 150~kHz (see Section~\ref{sec:SamplingTheorem}).
However, the higher rate was chosen, mainly because a higher number of signal samples beautify any plots that are created during analysis.
Since a thorough optimization of the DSP chain is not the focus of this thesis, this was an acceptable trade-off.\\

@@

At this point in the DSP chain, the signal resembles the frequency spectrum that is displayed in Figure~\ref{fig:fm_channel_baseband_freqs}.\\

@@

The receiver is implemented, following the block diagram in Figure~\ref{fig:bd_dsp_detailed}.
Further assumptions and simplifications, such as the fixed point arithmetic are described into detail above.

@@

An example of a plot that is created by the analysis is depicted in Figure~\ref{fig:matlab_analysis_time_domain}.
The diagram shows the respective signals in the time domain.
Above two timelines display the input data, as it is read from the input audio file.
The two diagrams in the center plot the Mono and LR Diff signals.
The bottom diagrams then represent the result after the entire signal processing chain.
It can be seen that the DSP chain is not able to demodulate the original signal without any error.
However, the result shows a clear separation of the left and right channels, as resembles the original input very closely, which is good enough for this case study.\\

@@

Further analysis is done in the frequency domain, especially in order to better understand the modulation of the different spectral parts in the FM channel.
An example thereof is shown in Figure~\ref{fig:matlab_analysis_freq_domain}.
The spectrum of an FM channel multiplex signal, as it is generated by a transmitter before the modulation, is shown in this diagram.
The spectrum after the demodulation through a receiver is shown in the lower diagram.
Spectral parts, such as the pilot tone at 19~kHz, or the LR Diff part around 38~kHz can be seen clearly.
The demodulated spectrum shows an artifact at 57~kHz, which does not exist in the original transmitter spectrum.
This is an artifact of the demodulation process, specifically the frequency-shift of the LR Diff part to baseband.
The artifact is very low in energy, because it is attenuated by a BP filter in the down-conversion process, and can thus be ignored for further processing, even though it can still be seen in the spectral analysis.

@@

GNU Radio provides a comprehensive set of DSP blocks, which are available to build various functionalities.
Multiple blocks are connected to a block diagram, that is very similar to the one in Figure~\ref{fig:bd_matlab_transmitter}.
The USRP device is used to broadcast the RF signal over the air.\\

@@

Special care needs to be taken when transmitting signals over the air.
The frequency spectrum is strongly limited by the radio regulations and may be monitored by the authorities.
Please make sure to adapt your setup to the respective local regulations, to avoid any legal issues.
Some of the regulations are mentioned in Section~\ref{sec:FrequencyBand}, but may differ in various countries, however.\\

\noindent
Figure~\ref{fig:bd_gnuradio_setup_transmitter} shows a block diagram of the transmitter application.

@@

This receiver works with both devices, the USRP b200mini and the RTL-SDR.
Figure~\ref{fig:bd_gnuradio_setup_receiver} shows a block diagram of the receiver application.

@@

Figure~\ref{fig:bd_cocotb_interface_simulator} gives an overview of how cocotb interfaces with a simulator.
The Python test cases are shown as a list of coroutines, which are managed by a scheduler.
This Python code interfaces with the simulator through standard interfaces like VPI.
The block called \textit{GPI} is developed by cocotb and abstracts the different simulator interfaces, such as VPI, VHPI or FLI.
In the simulator, the device under test (DUT) is instantiated.
The RTL simulation itself is managed by the simulators' scheduler.

@@

The architecture for this testbench is created under the assumptions and according to the guidelines that are developed in Chapter~\ref{cha:SystemArchitectureAndConcept}.\\

Figure~\ref{fig:bd_testbench_architecture_vhdl} shows a block diagram of the implemented architecture.
The split between the Python environment and the VHDL implementation is clearly visible but is represented from the developers perspective.
The actual simulator interfaces, as described in Section~\ref{sub:VHDL:FrameworkCocotb}, are hidden to maintain the overview of the implemented architecture.\\

@@

Figure~\ref{fig:register_engine} gives an overview of how the Register Engine script works.
The Python script takes the register definitions from the YAML file, and uses templates to generate the respecitive output files.\\

@@

A testbench is equally important in the HLS implementation, as it is for the VHDL implementation.
The HDL code for the IP is written in \cplusplus\ and thus the testbench is also written in this language.
The execution of the testbench is done in the Xilinx HLS software.
A detailed explanation of how an HLS testbench works in general is given in Chapter~\ref{cha:HLS}.\\

@@

The testbench is especially designed around the 'Design For Reuse' segment of the SIMILAR concept, which is introduced in Section~\ref{sec:SimilarProcess}.
This is the case because of two major parts.
%It is not only using the exact same Python code for analysis, but is also using parts of the application software code directly, to load the testbench input data.\\

@@

In the previous Section~\ref{sec:impl:hls:testbench} it is explained, that the HLS testbench generates output files during its execution.
These output files are analyzed with the exact same Python analysis code, that is already used in the VHDL testbench analysis.

@@

The testbench is developed in the language \cplusplus\ and does not have any restrictions in the usage thereof, as Section~\ref{sec:hls:workflow} described.
Thus, the \cplusplus\ code that is later used in the application software to interface with the IP, can already be used in the testbench here, for the same purpose.
This is especially useful in the development process of such a combined system, consisting of software and hardware, that relies on these two parts working together successfully.
The fact that the application software can directly be used in the testbench, enables the development of software that interfaces with the hardware in a very early stage of the project, before any actual hardware exists.
It also allows that the software part is already tested before the deployment on hardware.\\

Figure~\ref{fig:testbench_hls_overview_reuse} gives an overview of the testbench and specifically points out the reuse of the application software.
The diagram displays the HLS IP in the testbench and in the hardware with dotted lines, to emphasize that they are actually the same.\\

@@

HLS can automatically implement interfaces, as described in Section~\ref{sec:hls:coding:interfaces}.
This is a major feature that is used in the implementation of the HLS IP.\\

The design uses an AXI4-Lite memory mapped interface, as well as multiple AXI4-Streams.
Additionally, a data-only interface is used for the connected LEDs.
The actual implementation is demonstrated in Program~\ref{prog:hls_impl_interfaces}.
These few lines of code already implement the entire functionality of each interface type.\\

@@

%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Sample Rate Reduction}
\label{sec:impl:hls:sample_rate_reduction}

Multiple steps in the DSP chain require the operation of sample rate reduction.
This is explained in Section~\ref{sec:impl:dsp-chain}.\\

@@

The implementation of the sample rate reduction can be shown in a short code example in Program~\ref{prog:sample_rate_reduction_fifo_behaviour}.
The FM Demodulator runs in the oversampled domain and is executed \texttt{OSR\_RX} times in the inner for-loop.
Only the last sample is written to the \texttt{fm\_channel\_data} variable, which itself is an AXI stream type with an underlying FIFO depth of \texttt{OSR\_AUDIO}, as defined with the accompanying pragma.
Thus, the inner for-loop is executed \texttt{OSR\_AUDIO} times by the outer loop, to fill this FIFO until it is full.
The subsequent Channel Decoder then reads the entire FIFO and also implements downsampling with a similar behaviour.
Several samples are processed, where only the last one is forwarded to the subsequent processing step.
The FIFO behaviour of AXI stream variables is described in Section~\ref{sec:hls:coding:interfaces}.

### 2021-08-15 doc/thesis/chapters/introduction.tex
all /ref to ~/ref (to avoid newline separations)

Chapter \ref{cha:SignalProcessingTheory} describes the signal processing theory that is required to demodulate an FM broadcast radio signal.
This includes an introduction to receiver types, the description of an FM signal generally and also specifically as it is used in FM audio broadcasting, as well as various general topics of signal processing.
Chapter \ref{cha:SystemDesignProcess} covers the topic of a certain system design process, as well as system design tools that specifically target FPGAs.
Chapter \ref{cha:HLS} is about high-level synthesis, which includes a general introduction to the state of the art, as well as a description of the workflow, inclusively the coding and testbench features and characteristics.
Chapter \ref{cha:SystemArchitectureAndConcept} describes the chosen system architecture and concept for the project.
The actual implementation, with a detailed explanation of some selected components, is covered in Chapter \ref{cha:Implementation}.
Chapter \ref{cha:DeploymentOnHardware} gives an overview of the required parts of development that are necessary to deploy the deeloped system on actual hardware.
The comparison and achieved results are covered in Chapter \ref{cha:ComparisonAndResults}.
Chapter \ref{cha:ConclusionAndProspect} concludes this thesis with a retrospective and gives an outlook of potential future extensions to the project.

>>>>

Chapter~\ref{cha:SignalProcessingTheory} describes the signal processing theory that is required to demodulate an FM broadcast radio signal.
This includes an introduction to receiver types, the description of an FM signal generally and also specifically as it is used in FM audio broadcasting, as well as various general topics of signal processing.
Chapter~\ref{cha:SystemDesignProcess} covers the topic of a certain system design process, as well as system design tools that specifically target FPGAs.
Chapter~\ref{cha:HLS} is about high-level synthesis, which includes a general introduction to the state of the art, as well as a description of the workflow, inclusively the coding and testbench features and characteristics.
Chapter~\ref{cha:SystemArchitectureAndConcept} describes the chosen system architecture and concept for the project.
The actual implementation, with a detailed explanation of some selected components, is covered in Chapter~\ref{cha:Implementation}.
Chapter~\ref{cha:DeploymentOnHardware} gives an overview of the required parts of development that are necessary to deploy the deeloped system on actual hardware.
The comparison and achieved results are covered in Chapter~\ref{cha:ComparisonAndResults}.
Chapter~\ref{cha:ConclusionAndProspect} concludes this thesis with a retrospective and gives an outlook of potential future extensions to the project.

### 2021-08-15 doc/thesis/chapters/results.tex
remove unnecessary appendix
all /ref to ~/ref (to avoid newline separations)

The AXI stream interface however does not show the exact same behaviour.
Here, the HLS variant is using the AXI stream according to its protocol, while the VHDL variant is implemented differently.
It uses a simplified logic for the ready-flag, which reduces the effort in implementation.
However, from the perspective of the communicating blocks, the interface is usable like a regular AXI stream.
The implementation details are explained in Section \ref{sec:impl:vhdl:interfaces}.\\

@@

In the comparison against the Matlab model, the HLS variant achieves a very exact match of the output signal.
However, the VHDL implementation differs by a certain amount.
There seems to be an issue in the signal separation between the left and right audio channels.
This is visible in the analysis of the audio signal, as shown is Figure \ref{fig:audio_output_compare_tb_vs_hw}.
The signal strength of the left channel in the upper diagram is weaker than expected, while the right channel in the lower diagram is stronger.
This imbalance can also be observed by listening to the audio output via speakers.\\

@@

Here, the results of the simulation testbenches are compared against the values that are read from the FPGA directly.
Ideally, the values should be exactly the same.
However, Figure \ref{fig:audio_output_compare_tb_vs_hw} shows that this is not always the case.
Again, the HLS implementation does have a matching result, whereas the VHDL variant has deviating values.
The explanation here may be linked with above suspicions, but may also be linked to the initial reset values that exist in the hardware.
In the HLS implementation, the reset logic of several registers is implemented automatically, whereas in VHDL these conditions need to be implemented manually.
Therefore, the IPs internal state, i.e. the register values, may be different at the beginning, which leads to an error propagation throughout the design.
The entire DSP chain is built like a pipeline and therefor it takes a number of samples to process, before the 'wrong' intermediate values are flushed out.
All that may be the cause for the deviating values in the VHDL design.\\

@@

Several parts of the system design are taken into account to analyze the lines of code, to give an impression over the largest code parts in the project.
Figure \ref{fig:lines_of_code_pie_charts} shows multiple pie charts, which represent the proportions of the lines of code in the respective system design parts.
Table \ref{tab:lines_of_code} displays the actual numbers of code lines, to quantify the size of the respective parts.\\

@@

The metric of implementation time -- the time it takes to implement a specific feature -- is a subjective measure, since it heavily depends on the respective developer's experience, as already noted above.
In the present case, there is a strong background of knowledge and practical experience in VHDL design, while there is only very little of such in HLS design.
However, this metric is used in this section to specifically compare HLS with VHDL in this project.
Certain features are chosen for this comparison and their implementation times were logged during the actual implementation process.
The implementation is always considered as the HDL description, including the corresponding testbench code, to verify the respective correct functionality.
Table \ref{tab:implementation_time_logging} shows the direct comparison between HLS and VHDL implementation times for these features.\\

@@

The implementation of the sample decimation functionality differs slightly between the two variants, HLS and VHDL.
This is explained in Section \ref{sec:impl:hls:sample_rate_reduction}.
The implementation time of the HLS version is severely larger than the VHDL version.
This is mainly due to the limited knowledge of the functionality of the AXI stream interface at this stage in the project's implementation, just as previously mentioned with the FIR filter.
The initial version was implemented like the VHDL version -- a counter that only triggers the subsequent DSP blocks upon a certain number of samples was received.
However, this led to an endlessly blocking AXI stream in the testbench execution.
The reason was, that the read-function can only read a sample when the respective stream contains a sample, and that it is a blocking function.
Because the sample decimator did not write a sample every time, the subsequent read-function blocked indefinitely.
A thorough investigation in several user guides, in combination with step-by-step debugging, the issue was found and led to the current, working implementation, which processes a number of samples, before it forwards one sample.
The process of finding this issue explains the largely varying implementation times, as listed in Table \ref{tab:implementation_time_logging}.

@@

The implementation of the receiver structure requires the combination of several previously implemented blocks, such as the FIR filter and any interfaces.
This is done according to the Matlab model, to resemble its functionality.
The VHDL version is significantly more time consuming in comparison to HLS.
However, it is to note at this point that the VHDL version is implemented first, and thus some issues are already solved by the time of the HLS implementation.
Nevertheless, the VHDL implementation takes longer, as it requires more detailed care in the signal level.
This specifically includes the clock-alignment of any signals that need to be used in a combination.
For example, if two parallel FIR filters produce output signals that are to be added, they need to be aligned so that they are valid at the time of the summation.
This often requires analysis and debugging in a signal wave form.
In HLS this is not required, since the validity of a signal is determined internally by the HLS compiler, according to the usage of the respective data variables.
Another factor that adds up to the implementation time is the number of lines of code, as presented in Table \ref{tab:lines_of_code}.
The VHDL IP design requires 3000 lines, while the HLS IP design only takes 920 for the same functionality.
Again, this is a subjective measure -- several factors such as code style, or the developer's experience play a role -- but still shows a good representation of the implementation effort for this project.\\

@@

Figure \ref{fig:impl_time_analysis} shows this direct comparison of processed samples per second for the implementation variants.
The measurements always include the testbench setup-phase, and the run-phase.
The setup-phase mainly consists of loading the recorded data and its preparation, so that it can be fed into the DUT.
The run-phase then actually applies this test data to the DUT and stores the respective output results.
All time measurements are taken without the result analysis, since this is the exact same, shared code between all variants and thus takes the same amount of time everywhere.
Several measurements are done multiple times, to achieve an average result, that is independent of the current processor load of the host PC.\\

@@

Figure \ref{fig:hardware_utilization} shows the utilization numbers of the different FPGA resources.
It reveals that the manual VHDL implementation clearly requires more resources in the FPGA.
The study of the detailed utilization report however does not enable a deeper direct comparison of the respective single blocks, such as VHDL FIR filter against HLS FIR filter.
This is because the HLS compiler automatically applies optimizations, so that the direct connection between code and resource report is partially lost.
Examples for optimizations are inlining of functions, re-ordering of operations and time-sharing.
The latter has an especially large impact on the resource utilization.
Generally, in hardware design there is a trade-off between resource utilization and throughput, which leads to time-sharing.
For example, an entirely parallel-duplicated pipeline can process the double amount of data, but also requires the double amount of resources.
On the other hand, time-sharing a resource may take the double amount of time, thus only processes half of the data, but also only requires half of the resource utilization.
In HLS, this trade-off can be influenced and determined by the developer.
Depending on the requirements of the design, the optimization may either be set towards a lower resource utilization, or towards a high throughput.
In the current HLS implementation, no measures are taken towards any optimization and thus, the compiler is free to decide.
However, the only requirement is the correct functionality within the given situation, of course.
This specifically means that the throughput needs to be high enough, in order to process the audio signal output accordingly.\\

@@

Latency is the time it takes for the design to produce an output sample.
The latency of the HLS version is taken from the report that is produced after the C-synthesis of the HLS compiler.
The report shows an estimated resource utilization, as well as the latency as a number of clock cycles.
The latency is given with 2712 clock cycles.
For the VHDL version, a similar report is not available.
Instead, the testbench of the integration test is executed, which produces a signal wave form.
The latency is measured from this wave form and results in a number of around 2470 clock cycles.
With regards to latency, the two implementation variants appear to have a similar behaviour.

>>>>

The AXI stream interface however does not show the exact same behaviour.
Here, the HLS variant is using the AXI stream according to its protocol, while the VHDL variant is implemented differently.
It uses a simplified logic for the ready-flag, which reduces the effort in implementation.
However, from the perspective of the communicating blocks, the interface is usable like a regular AXI stream.
The implementation details are explained in Section~\ref{sec:impl:vhdl:interfaces}.\\

@@

In the comparison against the Matlab model, the HLS variant achieves a very exact match of the output signal.
However, the VHDL implementation differs by a certain amount.
There seems to be an issue in the signal separation between the left and right audio channels.
This is visible in the analysis of the audio signal, as shown is Figure~\ref{fig:audio_output_compare_tb_vs_hw}.
The signal strength of the left channel in the upper diagram is weaker than expected, while the right channel in the lower diagram is stronger.
This imbalance can also be observed by listening to the audio output via speakers.\\

@@

Here, the results of the simulation testbenches are compared against the values that are read from the FPGA directly.
Ideally, the values should be exactly the same.
However, Figure~\ref{fig:audio_output_compare_tb_vs_hw} shows that this is not always the case.
Again, the HLS implementation does have a matching result, whereas the VHDL variant has deviating values.
The explanation here may be linked with above suspicions, but may also be linked to the initial reset values that exist in the hardware.
In the HLS implementation, the reset logic of several registers is implemented automatically, whereas in VHDL these conditions need to be implemented manually.
Therefore, the IPs internal state, i.e. the register values, may be different at the beginning, which leads to an error propagation throughout the design.
The entire DSP chain is built like a pipeline and therefor it takes a number of samples to process, before the 'wrong' intermediate values are flushed out.
All that may be the cause for the deviating values in the VHDL design.\\

@@

Several parts of the system design are taken into account to analyze the lines of code, to give an impression over the largest code parts in the project.
Figure~\ref{fig:lines_of_code_pie_charts} shows multiple pie charts, which represent the proportions of the lines of code in the respective system design parts.
Table~\ref{tab:lines_of_code} displays the actual numbers of code lines, to quantify the size of the respective parts.\\

@@

The metric of implementation time -- the time it takes to implement a specific feature -- is a subjective measure, since it heavily depends on the respective developer's experience, as already noted above.
In the present case, there is a strong background of knowledge and practical experience in VHDL design, while there is only very little of such in HLS design.
However, this metric is used in this section to specifically compare HLS with VHDL in this project.
Certain features are chosen for this comparison and their implementation times were logged during the actual implementation process.
The implementation is always considered as the HDL description, including the corresponding testbench code, to verify the respective correct functionality.
Table~\ref{tab:implementation_time_logging} shows the direct comparison between HLS and VHDL implementation times for these features.\\

@@

The implementation of the sample decimation functionality differs slightly between the two variants, HLS and VHDL.
This is explained in Section~\ref{sec:impl:hls:sample_rate_reduction}.
The implementation time of the HLS version is severely larger than the VHDL version.
This is mainly due to the limited knowledge of the functionality of the AXI stream interface at this stage in the project's implementation, just as previously mentioned with the FIR filter.
The initial version was implemented like the VHDL version -- a counter that only triggers the subsequent DSP blocks upon a certain number of samples was received.
However, this led to an endlessly blocking AXI stream in the testbench execution.
The reason was, that the read-function can only read a sample when the respective stream contains a sample, and that it is a blocking function.
Because the sample decimator did not write a sample every time, the subsequent read-function blocked indefinitely.
A thorough investigation in several user guides, in combination with step-by-step debugging, the issue was found and led to the current, working implementation, which processes a number of samples, before it forwards one sample.
The process of finding this issue explains the largely varying implementation times, as listed in Table~\ref{tab:implementation_time_logging}.

@@

The implementation of the receiver structure requires the combination of several previously implemented blocks, such as the FIR filter and any interfaces.
This is done according to the Matlab model, to resemble its functionality.
The VHDL version is significantly more time consuming in comparison to HLS.
However, it is to note at this point that the VHDL version is implemented first, and thus some issues are already solved by the time of the HLS implementation.
Nevertheless, the VHDL implementation takes longer, as it requires more detailed care in the signal level.
This specifically includes the clock-alignment of any signals that need to be used in a combination.
For example, if two parallel FIR filters produce output signals that are to be added, they need to be aligned so that they are valid at the time of the summation.
This often requires analysis and debugging in a signal wave form.
In HLS this is not required, since the validity of a signal is determined internally by the HLS compiler, according to the usage of the respective data variables.
Another factor that adds up to the implementation time is the number of lines of code, as presented in Table~\ref{tab:lines_of_code}.
The VHDL IP design requires 3000 lines, while the HLS IP design only takes 920 for the same functionality.
Again, this is a subjective measure -- several factors such as code style, or the developer's experience play a role -- but still shows a good representation of the implementation effort for this project.\\

@@

Figure~\ref{fig:impl_time_analysis} shows this direct comparison of processed samples per second for the implementation variants.
The measurements always include the testbench setup-phase, and the run-phase.
The setup-phase mainly consists of loading the recorded data and its preparation, so that it can be fed into the DUT.
The run-phase then actually applies this test data to the DUT and stores the respective output results.
All time measurements are taken without the result analysis, since this is the exact same, shared code between all variants and thus takes the same amount of time everywhere.
Several measurements are done multiple times, to achieve an average result, that is independent of the current processor load of the host PC.\\

@@

Figure~\ref{fig:hardware_utilization} shows the utilization numbers of the different FPGA resources.
It reveals that the manual VHDL implementation clearly requires more resources in the FPGA.
The study of the detailed utilization report however does not enable a deeper direct comparison of the respective single blocks, such as VHDL FIR filter against HLS FIR filter.
This is because the HLS compiler automatically applies optimizations, so that the direct connection between code and resource report is partially lost.
Examples for optimizations are inlining of functions, re-ordering of operations and time-sharing.
The latter has an especially large impact on the resource utilization.
Generally, in hardware design there is a trade-off between resource utilization and throughput, which leads to time-sharing.
For example, an entirely parallel-duplicated pipeline can process the double amount of data, but also requires the double amount of resources.
On the other hand, time-sharing a resource may take the double amount of time, thus only processes half of the data, but also only requires half of the resource utilization.
In HLS, this trade-off can be influenced and determined by the developer.
Depending on the requirements of the design, the optimization may either be set towards a lower resource utilization, or towards a high throughput.
In the current HLS implementation, no measures are taken towards any optimization and thus, the compiler is free to decide.
However, the only requirement is the correct functionality within the given situation, of course.
This specifically means that the throughput needs to be high enough, in order to process the audio signal output accordingly.\\

@@

Latency is the time it takes for the design to produce an output sample.
The latency of the HLS version is taken from the report that is produced after the C-synthesis of the HLS compiler.
The report shows an estimated resource utilization, as well as the latency as a number of clock cycles.
The latency is given with 2712 clock cycles.
For the VHDL version, a similar report is not available.
Instead, the testbench of the integration test is executed, which produces a signal wave form.
The latency is measured from this wave form and results in a number of 2470 clock cycles.
With regards to latency, the two implementation variants appear to have a similar behaviour.
